DESING AND SIMULATION SOFTWARE FOR RHODOTRON TYPE ELECTRON ACCELERATORS

### 2024-01-07 msthesis.tex
Spiral corrected

\documentclass[a4paper,oneside,12pt]{report}
\usepackage[utf8x]{inputenc} % To use Unicode (e.g. Turkish) characters
\usepackage{amsmath, amsthm} % Some extra symbols
%\usepackage{subcaption}
\usepackage{styles/fbe_tez}
\usepackage[bottom]{footmisc}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{amssymb,bm,cite,graphicx, fixmath, texdraw}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{notoccite}
%\usepackage{subcaption}
\usepackage{subfigure}
%\usepackage{caption}
\usepackage{tabu}
\usepackage{minted}
\usepackage{xcolor}

@@

% LIST OF SYMBOLS PAGE
\begin{symbols}
    % The title will be typeset as "LIST OF SYMBOLS".
    % Use a separate \sym command for each symbols definition.
    % First, Latin symbols in alphabetical order
    \sym{$\vec{a}$}{Acceleration}
    \sym{\textbf{$\vec{B}$}}{Magnetic field}
    \sym{c}{Speed of light in vacuum}
    \sym{E}{Energy}
    \sym{$e^-$}{Electron}
    \sym{$e^- - \vec{E}$}{Electron Electric Field Interaction}
    \sym{$e^- - \vec{B}$}{Electron Magnetic Field Interaction}
    \sym{$e^- - EM$}{Electron Electro-Magnetic Field Interaction}
    \sym{\textbf{$\vec{E}$}}{Electric field}
    \sym{f}{Frequency}
    \sym{$\vec{F}$}{Force}
    \sym{$\vec{p}$}{Momentum}
    \sym{$\vec{r}$}{Position}
    \sym{t}{Time}
    \sym{T}{Period}
    \sym{$\vec{v}$}{Velocity}
    \sym{Z}{Shunt Impedance}
    % 1 EMPTY LINE BETWEEN LATIN AND GREEK SYMBOLS GROUP!!!
    \sym{}{}
    % Then Greek symbols in alphabetical order
    \sym{$\gamma$}{Lorentz Factor}
    \sym{$\beta$}{Beta Factor}
    \sym{$\phi_{lag}$}{Phase Lag}

@@

% LIST OF ABBREV PAGE
\begin{abbreviations}
    % Abbreviations in alphabetical order
   \sym{1D}{One Dimensional}
   \sym{2D}{Two Dimensional}
   \sym{3D}{Three Dimensional}
   \sym{eV}{Electron Volt}
   \sym{Hz}{Hertz}
   \sym{KAHVELab}{Kandilli Detector Accelerator and \\Instrumentation Laboratory}
   \sym{LF}{Leap-Frog}
   \sym{Linac}{Linear Particle Accelerator}
   \sym{ODE}{Ordinary Differential Equation}
   \sym{RF}{Radio Frequency}
   \sym{RK}{Runge-Kutta}
   \sym{RK4}{Forth Order Runge-Kutta}
   \sym{RMS}{Root Mean Square}
   \sym{POC}{Proof of Concept}
   \sym{CPU}{Central Processing Unit}
   \sym{s}{Second}
   \sym{T}{Tesla}
   \sym{V}{Volt}
   \sym{W}{Watt}

@@

\subsection{Relation between momentum and acceleration}
In classical mechanics, Newton's second law defines force, $\vecthreeBF{F}$, and relation
\vspace{-10pt}
\begin{equation} \label{eq:newton_second_law}
    \vecthreeBF{F} = \frac{d \vecthreeBF{p}}{d t} = m \frac{d \vecthreeBF{v}}{d t} = m \vecthreeBF{a}   ,
    \vspace{-10pt}
\end{equation}
where $\vecthreeBF{p}$ is the momentum, $m\vecthreeBF{v}$ of the particle.

@@

Considering these two statements, we can find the relation between momentum and acceleration as 
\vspace{-15pt}
\begin{eqnarray}
    \deriv{\vecthree{p}}{t} &=& m_0 \deriv{(\gamma \vecthree{v})}{t} = m_0 \{  \deriv{\gamma}{t}\vecthree{v} + \gamma \deriv{\vecthree{v}}{t}   \} ,\\
    \deriv{\gamma}{t} &=& \gamma^3 \dotProd{\beta}{\deriv{\beta}{t}} = \frac{\gamma^3}{c} \dotProd{\beta}{a}   , \\
    \deriv{\vecthree{p}}{t}  &=& m_0 \{   \frac{\gamma^3}{c} (\dotProd{\beta}{a} )\vecthree{v} + \gamma \deriv{\vecthree{v}}{t}  \} .
    \vspace{-10pt}
\end{eqnarray}
\vspace{-30pt}
\newline
Therefore force can be written as
\vspace{-15pt}
\begin{equation} \label{eq:Frel}
    \vecthreeBF{F}  = \gamma m_0 \{ \vecthreeBF{a} + \gamma^2(\dotProd{\beta}{\textbf{a}})\vecthree{\beta} \} .
    \vspace{-10pt}
\end{equation}
\clearpage
It is clear that acceleration is not necessarily parallel to the force. 
To start separating the parallel and perpendicular components relative to $\vecthree{\beta}$, we can find $\vecthreeBF{a}_{||}$ and $\vecthreeBF{F}_{||}$ as
\vspace{-10pt}
\begin{equation} \label{eq:a_F_parallels}
    \begin{aligned}
        \vecthreeBF{a}_{||} = \frac{(\dotProd{a}{\beta})}{\beta^2}\beta  ,
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{F}_{||} = \frac{(\dotProd{F}{\beta})}{\beta^2}\beta  ,
    \end{aligned}
    \vspace{-10pt}
\end{equation}
where 
\vspace{-10pt}
\begin{eqnarray}
    \dotProd{\textbf{F}}{\beta} &=& \gamma m_0 \{ \dotProd{\textbf{a}}{\beta} + \gamma^2(\dotProd{\beta}{\textbf{a}})\beta^2 \} \nonumber\\
                                &=& \gamma m_0 (\dotProd{\textbf{a}}{\beta}) \{ \gamma^2\beta^2  + 1\} .
                                \vspace{-10pt}
\end{eqnarray}
Then, $\gamma^2 \beta^2 + 1 = \gamma^2 $ and \fromeq{a_F_parallels} can be used to achieve
\vspace{-10pt}
\begin{eqnarray}
    \dotProd{\textbf{F}}{\beta} &=& m_0 \gamma^3 (\dotProd{\textbf{a}}{\beta} )   , \nonumber \\
    \vecthreeBF{F}_{||} &=& \frac{(\dotProd{\textbf{F}}{\beta})}{\beta^2} \vec{\beta} \nonumber \\
                        &=& m_0 \gamma^3 \frac{(\dotProd{\textbf{a}}{\beta} )}{\beta^2} \vec{\beta} \nonumber \\
                        &=& m_0 \gamma^3 \vecthreeBF{a}_{||} . \label{eq:Frel_parallel}
                        \vspace{-20pt}
\end{eqnarray}
Therefore from \fromeqs{Frel}{Frel_parallel}, force can be rearranged as
\vspace{-10pt}
\begin{eqnarray}
    \vecthreeBF{F}  &=& m_0 \gamma^3 \vecthreeBF{a}_{||} \beta^2 + m_0 \gamma \vecthreeBF{a} \nonumber\\
                    &=& m_0 \gamma^3 \vecthreeBF{a}_{||} \beta^2 + m_0 \gamma \{ \vecthreeBF{a}_{||} +  \vecthreeBF{a}_{\perp} \} \nonumber\\
                    &=& m_0 \vecthreeBF{a}_{||} \gamma \{ \gamma^2\beta^2 + 1 \} + m_0 \gamma \vecthreeBF{a}_{\perp} \nonumber\\
                    &=& m_0 \vecthreeBF{a}_{||} \gamma^3 + m_0 \gamma \vecthreeBF{a}_{\perp} \nonumber\\
                    &=& \vecthreeBF{F}_{||} + m_0 \gamma \vecthreeBF{a}_{\perp} .
                    \vspace{-10pt}
\end{eqnarray}

We finally have two separate equations which are in similar form with \fromeq{newton_second_law},
\vspace{-10pt}
\begin{equation} \label{eq:Frel_para_and_perp}
    \begin{aligned}
        \vecthreeBF{F}_{||} = \gamma^3  m_0 \vecthreeBF{a}_{||}  ,
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{F}_{\perp} = \gamma  m_0\vecthreeBF{a}_{\perp} .
    \end{aligned}
\end{equation}

\subsection{Lorentz Force} \label{sec:lorentz-force}
Force acting on a charged particle moving in electromagnetic fields is called Lorentz Force and is given by 
\vspace{-10pt}
\begin{equation} \label{eq:lf}
    \frac{d \vecthree{p}}{d t} = \vecthree{F}_L=q(\vecthree{E}+ \vecthree{v} \times \vecthree{B}) , 
    \vspace{-10pt}
\end{equation}
where the $q$ is the charge and $\vecthree{v}$ is the velocity of the particle.

\subsection{Relativistic Lorentz Force}
Similar to non-relativistic version, relativistic Lorentz Force is given by the 4-vector equality
\vspace{-10pt}
\begin{equation}
    \frac{d p^{\mu}}{d \tau} = q F^{\mu \nu} u_{\nu} ,
    \vspace{-10pt}
\end{equation}
where $ d \tau = d t / \gamma $, greek letters are Lorentz indices and \textit{p, F, u} are
\vspace{-10pt}
\begin{equation*}
    \begin{aligned}
        p^{\mu} = 
        \begin{bmatrix}
            W/c \\
            p_x \\
            p_y \\
            p_z
        \end{bmatrix}    
    \end{aligned} ,
    \qquad\qquad
    \begin{aligned}
        F^{\mu\nu} = 
        \begin{bmatrix}
                0       & -E_x/c   & -E_y/c    & -E_z/c \\
                E_x/c   &   0      & -B_z      & B_y    \\
                E_y/c   &   B_z    &  0        & -B_x   \\
                E_z/c   &   -B_y   & B_x       & 0   
        \end{bmatrix} 
    \end{aligned} ,
    \qquad\qquad
    \begin{aligned}
        u_{\nu} = \gamma
        \begin{bmatrix}
                 c \\
                -v_x \\
                -v_y \\
                -v_z \\
        \end{bmatrix} 
    \end{aligned} ,
    \vspace{-10pt}
\end{equation*}
where $W$ is the energy of the particle and $\gamma$ is the Lorentz factor mentioned in the previous section.

For $ \mu = 0 $, we have the time component of the equation as
\vspace{-10pt}
\begin{eqnarray}
    \frac{\gamma}{c}\frac{d W}{d t} &=& \frac{q \gamma \vecthree{E}\cdot \vecthree{v}}{c} = \frac{q \gamma}{c}  \frac{\vecthree{E}  \cdot d \vecthree{r}}{d t}   ,\\
    \frac{d W}{d t} &=& q \frac{\vecthree{E}\cdot d \vecthree{r}} {d t} .
    \vspace{-10pt}
\end{eqnarray}
This is the definition of work done by an electric field. For $ \mu = 1,2,3 $, we have the spacial components
\vspace{-10pt}
\begin{equation*}
    \frac{d \vecthree{p}}{d \tau} = \gamma \frac{d \vecthree{p}}{d t} = q \gamma (\vecthree{E} + \vecthree{v} \times \vecthree{B}) ,
    \vspace{-10pt}
\end{equation*}
which simplifies to non-relativistic Lorentz Force in \fromeq{lf}.

\subsection{Acceleration caused by Lorentz force}
Due to the nature of the cross product, Lorentz force caused by a magnetic field is always perpendicular to the velocity of the particle. 
Therefore the acceleration of the magnetic field is straightforward,
\vspace{-10pt}
\begin{equation*}
    \vecthreeBF{F}_B = \vecthreeBF{F}_{\perp} = \gamma  m_0\vecthreeBF{a}_{\perp} = \gamma  m_0\vecthreeBF{a}_{B} .
    \vspace{-10pt}
\end{equation*}
The same thing cannot be said about electric field however. 
It can create force in any direction with respect to velocity. Therefore, we have the equality
\vspace{-10pt}
\begin{equation} \label{eq:para_and_perp_acc_of_lorentz_force}
    \begin{aligned}
        \vecthreeBF{a}_{||} = \frac{q }{\gamma^3 m_0} \vecthreeBF{E}_{||}
    \end{aligned}  ,
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{a}_{\perp} = \frac{q}{\gamma m_0} \{ \vecthreeBF{E}_{\perp} + \crossProdBF{v}{B}   \} .
    \end{aligned}    
    \vspace{-10pt}
\end{equation}

Acceleration due to electric field can be simplified as
\vspace{-10pt}
\begin{eqnarray}
    \vecthreeBF{a}_{\{B=0\}} = \vecthreeBF{a}_{E}   &=&  \vecthreeBF{a}_{||} + \vecthreeBF{a}_{\perp \{B=0\}} \nonumber \\
                                &=& \frac{q}{m_0 \gamma} \{ \frac{\vecthreeBF{E}_{||}}{\gamma^2} + \vecthreeBF{E}_{\perp} \} \nonumber\\
                                &=& \frac{q}{m_0 \gamma} \{ \{1 - \beta^2\}\vecthreeBF{E}_{||} + \vecthreeBF{E}_{\perp} \} \nonumber\\
                                &=& \frac{q}{m_0 \gamma} \{ \vecthreeBF{E}_{||} + \vecthreeBF{E}_{\perp}-\beta^2\vecthreeBF{E}_{||} \} \nonumber\\
                                &=& \frac{q}{m_0 \gamma} \{ \vecthreeBF{E} - \beta^2\vecthreeBF{E}_{||} \} .
                                \vspace{-10pt}
\end{eqnarray}
Using the fact that $\vecthreeBF{E}_{||} = \vec{\beta}(\dotProd{\textbf{E}}{\beta})/\beta^2 $, we finally have
\vspace{-10pt}
\begin{equation} \label{eq:acc_of_E_and_B}
    \begin{aligned}
        \vecthreeBF{a}_{E} = \frac{q}{ \gamma m_0} \{ \vecthreeBF{E} - \vecthreeBF{v}\frac{(\dotProdBF{E}{v})}{c^2} \}
    \end{aligned}  ,
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{a}_{B} = \frac{q}{\gamma m_0} ( \crossProdBF{v}{B} ) ,
    \end{aligned}
    \vspace{-10pt}
\end{equation}
where $\vecthreeBF{a}_{E}$ consists of two components in the direction of $\vecthreeBF{E}$ and $\vecthreeBF{v}$.
The component in the direction of $\vecthreeBF{v}$ depends on the $\dotProdBF{E}{v}$, 
resulting in $\vecthreeBF{a}_{E} || \vecthreeBF{E}$ if $\vecthreeBF{E} \perp \vecthreeBF{v}$.

@@

As particles gain energy in the accelerator, they approach relativistic speeds, where relativistic effects become significant. 
Special relativity governs the increase in mass and energy of the particles as they approach the speed of light, providing insights into the behavior of matter at high energies.
% TODO : ADD CIRCULAR AND LINEAR ACCELERATOR PICTURES HERE
\begin{figure}[H]
    \centering
    \includegraphics[width=.9\textwidth]{./figures/kahvelab_linac.png}
    \vspace{20pt}
    \caption{A linear proton accelerator in KAHVELab \cite{kahvelab_linac}.}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=.75\textwidth]{./figures/pill_box.jpeg}
    \vspace{20pt}
    \caption{An RF cavity used in KAHVELab.}
\end{figure}

In addition to accelerating the particles, RF cavities are often designed to provide focusing forces. 
By carefully shaping the cavity and adjusting the electromagnetic fields, the particles can experience focusing effects as they pass through the cavity. 
This helps to maintain a tight and controlled beam. To ensure efficient acceleration, it is essential to maintain phase stability. 
This means that the particles should experience the strongest electric fields at the correct time during their passage through the cavity. 
Precise timing and synchronization of the RF power source with the particle beam are crucial to achieve phase stability and maximize energy transfer.

@@

\subsubsection{Resonance Frequency}
Resonance frequency is the frequency in which the electromagnetic fields form standing waves in a cavity.
It is determined by the physical dimensions and the speed of light in the cavity's medium. 
In a rectangular cavity for example, the resonance frequency can be calculated by
\vspace{-10pt}\begin{equation}
    f_{klm} = \frac{c}{2\pi \sqrt{\mu_r \epsilon_r}} \sqrt{(\frac{k}{w})^2 + (\frac{l}{u})^2 + (\frac{m}{v})^2}  ,
    \label{eq:f_r_rec}
\vspace{-10pt}\end{equation}
where $w$, $u$, $v$ are the dimensions of the cavity, $\mu_r$ and $\epsilon_r$ are relative permability and permittivity of the cavity respectably.

\subsubsection{Bunch}
Bunch refers to a tightly grouped collection of charged particles, such as electrons or protons, that are accelerated and maintained close together within a particle accelerator.

\subsubsection{Phase Stability}
Phase stability refers to the preservation of the timing relationship between particles and fields within an accelerator. 
It ensures that the phases of various electromagnetic fields or particles remain synchronized, 
which is crucial for achieving efficient particle acceleration. 
Maintaining phase stability is essential to prevent particles from becoming out of phase as they travel through accelerator structures, 
ensuring that they receive the correct energy boosts and interact predictably with detectors. 
Deviations in phase stability can lead to particle loss and decreased beam quality.

\subsubsection{Phase Lag}
Phase lag refers to the time delay between the oscillations of two interacting waveforms or particles. 
It describes the difference in phase angles within their respective cycles, between two signals.

>>>>

\documentclass[a4paper,oneside,12pt]{report}
\usepackage[utf8x]{inputenc} % To use Unicode (e.g. Turkish) characters
\usepackage{amsmath, amsthm} % Some extra symbols
%\usepackage{subcaption}
\usepackage{styles/fbe_tez}
\usepackage[bottom]{footmisc}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{multirow}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{amssymb,bm,cite,graphicx, fixmath, texdraw}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{notoccite}
%\usepackage{subcaption}
\usepackage{subfigure}
%\usepackage{caption}
\usepackage{tabu}
\usepackage{minted}
\usepackage{xcolor}
\usepackage{showframe}

@@

% LIST OF SYMBOLS PAGE
\begin{symbols}
    % The title will be typeset as "LIST OF SYMBOLS".
    % Use a separate \sym command for each symbols definition.
    % First, Latin symbols in alphabetical order
    \sym{$\vec{a}$}{Acceleration}
    \sym{\textbf{$\vec{B}$}}{Magnetic field}
    \sym{c}{Speed of light in vacuum}
    \sym{E}{Energy}
    \sym{$e^-$}{Electron}
    \sym{$e^- - \vec{E}$}{Electron Electric Field Interaction}
    \sym{$e^- - \vec{B}$}{Electron Magnetic Field Interaction}
    \sym{$e^- - EM$}{Electron Electro-Magnetic Field Interaction}
    \sym{\textbf{$\vec{E}$}}{Electric field}
    \sym{f}{Frequency}
    \sym{$\vec{F}$}{Force}
    \sym{$\vec{p}$}{Momentum}
    \sym{$\vec{r}$}{Position}
    \sym{t}{Time}
    \sym{T}{Period}
    \sym{$\vec{v}$}{Velocity}
    \sym{Z}{Shunt Impedance}
    % 1 EMPTY LINE BETWEEN LATIN AND GREEK SYMBOLS GROUP!!!
    \sym{}{}
    % Then Greek symbols in alphabetical order
    \sym{$\beta$}{Beta Factor}
    \sym{$\gamma$}{Lorentz Factor}
    \sym{$\phi_{lag}$}{Phase Lag}

@@

% LIST OF ABBREV PAGE
\begin{abbreviations}
    % Abbreviations in alphabetical order
   \sym{1D}{One Dimensional}
   \sym{2D}{Two Dimensional}
   \sym{3D}{Three Dimensional}
   \sym{eV}{Electron Volt}
   \sym{Hz}{Hertz}
   \sym{KAHVELab}{Kandilli Detector Accelerator and \\Instrumentation Laboratory}
   \vspace{10pt}
   \sym{LF}{Leap-Frog}
   \sym{Linac}{Linear Particle Accelerator}
   \sym{ODE}{Ordinary Differential Equation}
   \sym{RF}{Radio Frequency}
   \sym{RK}{Runge-Kutta}
   \sym{RK4}{Forth Order Runge-Kutta}
   \sym{RMS}{Root Mean Square}
   \sym{POC}{Proof of Concept}
   \sym{CPU}{Central Processing Unit}
   \sym{s}{Second}
   \sym{T}{Tesla}
   \sym{V}{Volt}
   \sym{W}{Watt}

@@

\subsection{Relation Between Momentum and Acceleration}
In classical mechanics, Newton's second law defines force, $\vecthreeBF{F}$, and relation
\vspace{-10pt}
\begin{equation} \label{eq:newton_second_law}
    \vecthreeBF{F} = \frac{d \vecthreeBF{p}}{d t} = m \frac{d \vecthreeBF{v}}{d t} = m \vecthreeBF{a}   ,
    \vspace{-10pt}
\end{equation}
where $\vecthreeBF{p}$ is the momentum, $m\vecthreeBF{v}$ of the particle.

@@

Considering these two statements, we can find the relation between momentum and acceleration as 
\vspace{-15pt}
\begin{eqnarray}
    \deriv{\vecthree{p}}{t} &=& m_0 \deriv{(\gamma \vecthree{v})}{t} = m_0 \{  \deriv{\gamma}{t}\vecthree{v} + \gamma \deriv{\vecthree{v}}{t}   \} ,\\
    \deriv{\gamma}{t} &=& \gamma^3 \dotProd{\beta}{\deriv{\beta}{t}} = \frac{\gamma^3}{c} \dotProd{\beta}{a}   , \\
    \deriv{\vecthree{p}}{t}  &=& m_0 \{   \frac{\gamma^3}{c} (\dotProd{\beta}{a} )\vecthree{v} + \gamma \deriv{\vecthree{v}}{t}  \} .
    \vspace{-15pt}
\end{eqnarray}
\vspace{-35pt}
\newline
Therefore force can be written as
\vspace{-20pt}
\begin{equation} \label{eq:Frel}
    \vecthreeBF{F}  = \gamma m_0 \{ \vecthreeBF{a} + \gamma^2(\dotProd{\beta}{\textbf{a}})\vecthree{\beta} \} .
    \vspace{-15pt}
\end{equation}

It is clear that acceleration is not necessarily parallel to the force. 
To start separating the parallel and perpendicular components relative to $\vecthree{\beta}$, we can find $\vecthreeBF{a}_{||}$ and $\vecthreeBF{F}_{||}$ as
\vspace{-10pt}
\begin{equation} \label{eq:a_F_parallels}
    \begin{aligned}
        \vecthreeBF{a}_{||} = \frac{(\dotProd{a}{\beta})}{\beta^2}\beta  ,
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{F}_{||} = \frac{(\dotProd{F}{\beta})}{\beta^2}\beta  ,
    \end{aligned}
    \vspace{-10pt}
\end{equation}
where 
\vspace{-17pt}
\begin{eqnarray}
    \dotProd{\textbf{F}}{\beta} &=& \gamma m_0 \{ \dotProd{\textbf{a}}{\beta} + \gamma^2(\dotProd{\beta}{\textbf{a}})\beta^2 \} \nonumber\\
                                &=& \gamma m_0 (\dotProd{\textbf{a}}{\beta}) \{ \gamma^2\beta^2  + 1\} .
\end{eqnarray}

\vspace{-40pt}
\noindent
Then, $\gamma^2 \beta^2 + 1 = \gamma^2 $ and \fromeq{a_F_parallels} can be used to achieve
\vspace{-20pt}
\begin{eqnarray}
    \dotProd{\textbf{F}}{\beta} &=& m_0 \gamma^3 (\dotProd{\textbf{a}}{\beta} )   , \nonumber \\
    \vecthreeBF{F}_{||} &=& \frac{(\dotProd{\textbf{F}}{\beta})}{\beta^2} \vec{\beta} \nonumber \\
                        &=& m_0 \gamma^3 \frac{(\dotProd{\textbf{a}}{\beta} )}{\beta^2} \vec{\beta} \nonumber \\
                        &=& m_0 \gamma^3 \vecthreeBF{a}_{||} . \label{eq:Frel_parallel}
\end{eqnarray}

\vspace{-41pt}
\noindent
Therefore from \fromeqs{Frel}{Frel_parallel}, force can be rearranged as
\vspace{-20pt}
\begin{eqnarray}
    \vecthreeBF{F}  &=& m_0 \gamma^3 \vecthreeBF{a}_{||} \beta^2 + m_0 \gamma \vecthreeBF{a} \nonumber\\
                    &=& m_0 \gamma^3 \vecthreeBF{a}_{||} \beta^2 + m_0 \gamma \{ \vecthreeBF{a}_{||} +  \vecthreeBF{a}_{\perp} \} \nonumber\\
                    &=& m_0 \vecthreeBF{a}_{||} \gamma \{ \gamma^2\beta^2 + 1 \} + m_0 \gamma \vecthreeBF{a}_{\perp} \nonumber\\
                    &=& m_0 \vecthreeBF{a}_{||} \gamma^3 + m_0 \gamma \vecthreeBF{a}_{\perp} \nonumber\\
                    &=& \vecthreeBF{F}_{||} + m_0 \gamma \vecthreeBF{a}_{\perp} .
\end{eqnarray}

\vspace{-20pt}
We finally have two separate equations which are in similar form with \fromeq{newton_second_law},
\vspace{-20pt}
\begin{equation} \label{eq:Frel_para_and_perp}
    \begin{aligned}
        \vecthreeBF{F}_{||} = \gamma^3  m_0 \vecthreeBF{a}_{||}  ,
    \end{aligned}
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{F}_{\perp} = \gamma  m_0\vecthreeBF{a}_{\perp} .
    \end{aligned}
    \vspace{-18pt}
\end{equation}

\subsection{Lorentz Force} \label{sec:lorentz-force}
Force acting on a charged particle moving in electromagnetic fields is called Lorentz Force and is given by 
\vspace{-12pt}
\begin{equation} \label{eq:lf}
    \frac{d \vecthree{p}}{d t} = \vecthree{F}_L=q(\vecthree{E}+ \vecthree{v} \times \vecthree{B}) , 
    \vspace{-12pt}
\end{equation}
where the $q$ is the charge and $\vecthree{v}$ is the velocity of the particle.

\subsection{Relativistic Lorentz Force}
Similar to non-relativistic version, relativistic Lorentz Force is given by the 4-vector equality
\vspace{-11pt}
\begin{equation}
    \frac{d p^{\mu}}{d \tau} = q F^{\mu \nu} u_{\nu} ,
    \vspace{-11pt}
\end{equation}
where $ d \tau = d t / \gamma $, greek letters are Lorentz indices and \textit{p, F, u} are
\vspace{-10pt}
\begin{equation}
    \begin{aligned}
        p^{\mu} = 
        \begin{bmatrix}
            W/c \\
            p_x \\
            p_y \\
            p_z
        \end{bmatrix}    
    \end{aligned} ,
    \quad
    \begin{aligned}
        F^{\mu\nu} = 
        \begin{bmatrix}
                0       & -E_x/c   & -E_y/c    & -E_z/c \\
                E_x/c   &   0      & -B_z      & B_y    \\
                E_y/c   &   B_z    &  0        & -B_x   \\
                E_z/c   &   -B_y   & B_x       & 0   
        \end{bmatrix} 
    \end{aligned} ,
    \quad
    \begin{aligned}
        u_{\nu} = \gamma
        \begin{bmatrix}
                 c \\
                -v_x \\
                -v_y \\
                -v_z \\
        \end{bmatrix} 
    \end{aligned} ,
    \vspace{-10pt}
\end{equation}
where $W$ is the energy of the particle and $\gamma$ is the Lorentz factor mentioned in the previous section.

For $ \mu = 0 $, we have the time component of the equation as
\vspace{-12pt}
\begin{eqnarray}
    \frac{\gamma}{c}\frac{d W}{d t} &=& \frac{q \gamma \vecthree{E}\cdot \vecthree{v}}{c} = \frac{q \gamma}{c}  \frac{\vecthree{E}  \cdot d \vecthree{r}}{d t}   ,\\
    \frac{d W}{d t} &=& q \frac{\vecthree{E}\cdot d \vecthree{r}} {d t} .
\end{eqnarray}

\vspace{-33pt}
\noindent
This is the definition of work done by an electric field. For $ \mu = 1,2,3 $, we have the spacial components
\vspace{-11pt}
\begin{equation}
    \frac{d \vecthree{p}}{d \tau} = \gamma \frac{d \vecthree{p}}{d t} = q \gamma (\vecthree{E} + \vecthree{v} \times \vecthree{B}) ,
    \vspace{-11pt}
\end{equation}
which simplifies to non-relativistic Lorentz Force in \fromeq{lf}.

\subsection{Acceleration Caused by Lorentz Force}
Due to the nature of the cross product, Lorentz force caused by a magnetic field is always perpendicular to the velocity of the particle. 
Therefore the acceleration of the magnetic field is straightforward,
\vspace{-10pt}
\begin{equation}
    \vecthreeBF{F}_B = \vecthreeBF{F}_{\perp} = \gamma  m_0\vecthreeBF{a}_{\perp} = \gamma  m_0\vecthreeBF{a}_{B} .
    \vspace{-20pt}
\end{equation}
The same thing cannot be said about electric field however. 
It can create force in any direction with respect to velocity. Therefore, we have the equality
\vspace{-15pt}
\begin{equation} \label{eq:para_and_perp_acc_of_lorentz_force}
    \begin{aligned}
        \vecthreeBF{a}_{||} = \frac{q }{\gamma^3 m_0} \vecthreeBF{E}_{||}
    \end{aligned}  ,
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{a}_{\perp} = \frac{q}{\gamma m_0} \{ \vecthreeBF{E}_{\perp} + \crossProdBF{v}{B}   \} .
    \end{aligned}    
    \vspace{-10pt}
\end{equation}

Acceleration due to electric field can be simplified as
\vspace{-20pt}
\begin{eqnarray}
    \vecthreeBF{a}_{\{B=0\}} = \vecthreeBF{a}_{E}   &=&  \vecthreeBF{a}_{||} + \vecthreeBF{a}_{\perp \{B=0\}} \nonumber \\
                                &=& \frac{q}{m_0 \gamma} \{ \frac{\vecthreeBF{E}_{||}}{\gamma^2} + \vecthreeBF{E}_{\perp} \} \nonumber\\
                                &=& \frac{q}{m_0 \gamma} \{ \{1 - \beta^2\}\vecthreeBF{E}_{||} + \vecthreeBF{E}_{\perp} \} \nonumber\\
                                &=& \frac{q}{m_0 \gamma} \{ \vecthreeBF{E}_{||} + \vecthreeBF{E}_{\perp}-\beta^2\vecthreeBF{E}_{||} \} \nonumber\\
                                &=& \frac{q}{m_0 \gamma} \{ \vecthreeBF{E} - \beta^2\vecthreeBF{E}_{||} \} .
\end{eqnarray}

\vspace{-27pt}
\noindent
Using the fact that $\vecthreeBF{E}_{||} = \vec{\beta}(\dotProd{\textbf{E}}{\beta})/\beta^2 $, we finally have
\vspace{-10pt}
\begin{equation} \label{eq:acc_of_E_and_B}
    \begin{aligned}
        \vecthreeBF{a}_{E} = \frac{q}{ \gamma m_0} \{ \vecthreeBF{E} - \vecthreeBF{v}\frac{(\dotProdBF{E}{v})}{c^2} \}
    \end{aligned}  ,
    \qquad \qquad
    \begin{aligned}
        \vecthreeBF{a}_{B} = \frac{q}{\gamma m_0} ( \crossProdBF{v}{B} ) ,
    \end{aligned}
    \vspace{-12pt}
\end{equation}
where $\vecthreeBF{a}_{E}$ consists of two components in the direction of $\vecthreeBF{E}$ and $\vecthreeBF{v}$.
The component in the direction of $\vecthreeBF{v}$ depends on the $\dotProdBF{E}{v}$, 
resulting in $\vecthreeBF{a}_{E} || \vecthreeBF{E}$ if $\vecthreeBF{E} \perp \vecthreeBF{v}$.

@@

As particles gain energy in the accelerator, they approach relativistic speeds, where relativistic effects become significant. 
Special relativity governs the increase in mass and energy of the particles as they approach the speed of light, providing insights into the behavior of matter at high energies.

\vspace{22pt}
\begin{figure}[H]
    \centering
    \includegraphics[width=.7\textwidth]{./figures/kahvelab_linac.png}
    %\vspace{22pt}
    \vspace{-1pt}
    \caption{A linear proton accelerator in KAHVELab \cite{kahvelab_linac}.}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=.5\textwidth]{./figures/pill_box.jpeg}
    %\vspace{22pt}
    \vspace{-1pt}
    \caption{An RF cavity used in KAHVELab.}
\end{figure}

\vspace{-15pt}
In addition to accelerating the particles, RF cavities are often designed to provide focusing forces. 
By carefully shaping the cavity and adjusting the electromagnetic fields, the particles can experience focusing effects as they pass through the cavity. 
This helps to maintain a tight and controlled beam. To ensure efficient acceleration, it is essential to maintain phase stability. 
This means that the particles should experience the strongest electric fields at the correct time during their passage through the cavity. 
Precise timing and synchronization of the RF power source with the particle beam are crucial to achieve phase stability and maximize energy transfer.

@@

\subsubsection{\hspace{4pt}Resonance Frequency}\hspace{4pt}
Resonance frequency is the frequency in which the electromagnetic fields form standing waves in a cavity.
It is determined by the physical dimensions and the speed of light in the cavity's medium. 
In a rectangular cavity for example, the resonance frequency can be calculated by
\vspace{-10pt}\begin{equation}
    f_{klm} = \frac{c}{2\pi \sqrt{\mu_r \epsilon_r}} \sqrt{(\frac{k}{w})^2 + (\frac{l}{u})^2 + (\frac{m}{v})^2}  ,
    \label{eq:f_r_rec}
\vspace{-10pt}\end{equation}
where $w$, $u$, $v$ are the dimensions of the cavity, $\mu_r$ and $\epsilon_r$ are relative permability and permittivity of the cavity respectably.

\subsubsection{\hspace{4pt}Bunch}\hspace{4pt}
Bunch refers to a tightly grouped collection of charged particles, such as electrons or protons, that are accelerated and maintained close together within a particle accelerator.

\subsubsection{\hspace{4pt}Phase Stability}\hspace{4pt}
Phase stability refers to the preservation of the timing relationship between particles and fields within an accelerator. 
It ensures that the phases of various electromagnetic fields or particles remain synchronized, 
which is crucial for achieving efficient particle acceleration. 
Maintaining phase stability is essential to prevent particles from becoming out of phase as they travel through accelerator structures, 
ensuring that they receive the correct energy boosts and interact predictably with detectors. 
Deviations in phase stability can lead to particle loss and decreased beam quality.

\subsubsection{\hspace{4pt}Phase Lag}\hspace{4pt}
Phase lag refers to the time delay between the oscillations of two interacting waveforms or particles. 
It describes the difference in phase angles within their respective cycles, between two signals.
